Supplemental material for `Eliminating object prior-bias from sparse-projection tomographic reconstructions'

### 2021-10-12 paper/post_tci.tex
changed the section on tuning and corresponding changes to supplementary

There also exist methods for estimating new changes (in the test)
directly in projection space. In such techniques which can be used in
longitudinal studies, the object-prior is projected in the same set of
views as the current set of test measurements, and a set of projection
differences is computed. These projection differences are essentially
tomographic projections of differences between the unknown test and
the known object-prior. Hence, a variety of reconstruction algorithms
can be used to estimate such a difference image which can then be
added to the object-prior to yield the final estimate of the
test. Such methods have been proposed
in~\cite{Pourmorteza2015,Lee2012}. However, reconstruction of the
difference image will inherently contain sub-sampling artefacts, and
these artefacts will appear in the final reconstruction as there is no
mechanism to mitigate them (unlike the method that we propose). This
can be seen in Sec.~3 our supplementary material~\cite{supp_paper},
where we have compared our method with~\cite{Lee2012}.

@@

$\lambda_1$, $\lambda_2$ are tunable weights given to the
TV-prior and object-prior terms respectively. The unknowns
$\boldsymbol{x}$ and $\boldsymbol{\alpha}$ are solved by alternately
minimizing $J_{\boldsymbol{\alpha}}(\boldsymbol{x})$ using a fixed
$\boldsymbol{\alpha}$, and $J_{\boldsymbol{x}}(\boldsymbol{\alpha})$
using the resultant $\boldsymbol{x}$.
$J_{\boldsymbol{\alpha}}(\boldsymbol{x})$ is solved for using an
optimal first-order method for large-scale TV regularization presented
in~\cite{TVReg} and available in~\cite{TVReg-lib}.

@@

\section{Discussion}
\label{sec:discussion}
In this section, we discuss tuning of parameters, computational complexity of our method, and requirements for applying deep learning techniques in longitudinal few-view reconstruction.
\subsection{Tuning of parameters}
$\lambda_1$, $\lambda_2$ and $k$ are the three
hyper-parameters that need to be chosen carefully for optimal
reconstruction. In our experiments, $\lambda_1$ was tuned to maximize
SSIM of the whole reconstructed test volume for TV reconstruction of
each dataset. This value was retained for spatially-varying
prior-based reconstruction as well. The value of $\lambda_2$ largely
depends on the amount of artefacts we aim to remove by using
object-prior at the cost of their dominance in the new regions. This
value was chosen to lie between $0-3$ for our datasets. Finally, the
hyper-parameter $k$ defines the sensitivity of the weights map to the
difference between test image and object-prior (projection of
test onto the space of object-prior). When $k=0$, our method 
behaves similar to the reconstruction using TV-only regularization. As $k$ increases, the weights map starts
capturing new changes in the test, at the cost of detecting a few
false positives i.e., false new changes. In other words, as the
weights-map becomes more sensitive to the difference between the test
and object-prior, it becomes more noisy. In order to visualize the
effect of the hyper-parameter $k$, we performed 2D reconstructions on
okra dataset for different values of
$k$. Fig.~\ref{fig:few_view_okra_2D_weights} shows the weights map
obtained for each of the $k$ values and
Fig.~\ref{fig:reconstructions_as_k_varies} shows the corresponding
final reconstructions. We run our experiments for a few values of $k$
and choose the best reconstruction based on our tolerance for noise
in the weights-map.
When a large number of templates are available, this parameter can be
chosen by assuming one of the templates as test and choosing the $k$
that gave the best reconstruction. Alternatively, in cases where one
wishes to completely avoid the use of this hyper-parameter, one can
construct a binary weights-map using a learning based method described
in the supplementary material~\cite{supp_paper}.

>>>>

There also exist methods for estimating new changes (in the test)
directly in projection space. In such techniques which can be used in
longitudinal studies, the object-prior is projected in the same set of
views as the current set of test measurements, and a set of projection
differences is computed. These projection differences are essentially
tomographic projections of differences between the unknown test and
the known object-prior. Hence, a variety of reconstruction algorithms
can be used to estimate such a difference image which can then be
added to the object-prior to yield the final estimate of the
test. Such methods have been proposed
in~\cite{Pourmorteza2015,Lee2012}. However, reconstruction of the
difference image will inherently contain sub-sampling artefacts, and
these artefacts will appear in the final reconstruction as there is no
mechanism to mitigate them (unlike the method that we propose). This
can be seen in Sec.~4 of our supplementary material~\cite{supp_paper},
where we have compared our method with~\cite{Lee2012}.

@@

$\lambda_1$ and  $\lambda_2$ are tunable weights given to the
TV-prior and object-prior terms respectively. The unknowns
$\boldsymbol{x}$ and $\boldsymbol{\alpha}$ are solved by alternately
minimizing $J_{\boldsymbol{\alpha}}(\boldsymbol{x})$ using a fixed
$\boldsymbol{\alpha}$, and $J_{\boldsymbol{x}}(\boldsymbol{\alpha})$
using the resultant $\boldsymbol{x}$.
$J_{\boldsymbol{\alpha}}(\boldsymbol{x})$ is solved for using an
optimal first-order method for large-scale TV regularization presented
in~\cite{TVReg} and available in~\cite{TVReg-lib}.

@@

\section{Discussion}
\label{sec:discussion}
In this section, we discuss tuning of parameters, computational
complexity of our method, and requirements for applying deep learning
techniques in longitudinal few-view reconstruction.

\subsection{Tuning of parameters}
$\lambda_1$, $\lambda_2$ and $k$ are the three hyper-parameters that
need to be chosen carefully for optimal reconstruction. In our
experiments, $\lambda_1$ was tuned in such a way as to maximize the
SSIM of the complete reconstructed test volume using the TV-only
regularizer.  In particular, the RoI was not considered.  Now, this
value was retained for our spatially-varying prior-based
reconstruction as well.  In principle, one could have jointly
optimized both $\lambda_1$ and $\lambda_2$ to get even better results
compared to the one we have in Section~\ref{results}.  The value of
$\lambda_2$ largely depends on the amount of artefacts one aims to
remove by using object-prior at the cost of their dominance in the new
regions. This value was chosen to lie between $0-3$ for our datasets
(having the fixed $\lambda_1$).

The hyper-parameter $k$ defines the sensitivity of the weights map to
the difference between test image and object-prior (projection of test
onto the space of object-prior). When $k=0$, our method behaves
similar to the reconstruction using TV-only regularization. As $k$
increases, the weights map starts capturing new changes in the test,
at the cost of detecting a few false positives i.e., false new
changes. In other words, as the weights-map becomes more sensitive to
the difference between the test and object-prior, it becomes more
noisy. In order to visualize the effect of the hyper-parameter $k$, we
performed 2D reconstructions on okra dataset for different values of
$k$. Fig.~\ref{fig:few_view_okra_2D_weights} shows the weights map
obtained for each of the $k$ values and
Fig.~\ref{fig:reconstructions_as_k_varies} shows the corresponding
final reconstructions. We run our experiments for a few values of $k$
and empirically choose the best reconstruction based on our tolerance
for noise in the weights-map.  However, when a large number of
templates are available, instead of this empirical method, this
parameter can be chosen by assuming one of the templates as a
pseudo-test and validating the $k$ that results in the best
reconstruction (also see Sec~3 of \cite{supp_paper}.

Alternatively, in cases where one wishes to completely avoid the use
of this hyper-parameter, one can construct a binary weights-map using
a learning based method described in the supplementary
material~\cite{supp_paper}.

### 2021-10-12 paper/post_tci_supplementary.tex
changed the section on tuning and corresponding changes to supplementary

\usepackage{boxedminipage}
%\usepackage{geometry}
\usepackage[colorlinks=true]{hyperref}
\title{Supplemental material for `Eliminating object prior-bias from sparse-projection
tomographic reconstructions'}
\begin{document}
\maketitle
This supplementary material consists of details which we did not include in the main paper for purposes of brevity. This document consists of links to view our complete 3D reconstructions, more details on why we use many eigenspaces, an alternate method to compute weights map and comparison with a method in literature.
\tableofcontents

@@

\section{Links to view our 3D reconstructions}
Here we present the 3D reconstruction results discussed in Sec.~6 of the main paper. For each dataset, the links contain 2 videos: one showing all the templates and test volumes, and the other showing the reconstructed volumes by various methods. The videos are in .avi format and display reconstructions in a slice-by-slice manner.

@@

%------------------------------------------------------
\section{Motivation for using multiple eigenspaces in our method}

Here we present a schematic explaining the advantage of using multiple eigenspaces within the
proposed spatially-varying reconstruction technique.\\

@@

%\clearpage
%----------------------------------------------------------
\section{An alternate method to compute weights -map}
The value of $k$ (discussed in Sec.~7 of the main paper) is typically dependent on the number of false positives the system can allow in exchange for the detection of more true positives (regions of new changes). It is this choice and not the particular type of dataset that affects the selection of $k$. In our work, the chosen $k$ corresponds to the one that gives minimal tolerance to false positives. However, in case one wishes to completely avoid the use of this hyper-parameter, one can construct a binary weights-map using an alternate method described below.

\begin{enumerate}
\item For the purposes of training, treat one of the previously scanned images as test. Annotate the region of differences between the test and all of the previously scanned images based on domain knowledge.
\item Compute the difference between the pilot reconstruction of the test and its projection onto the eigenspace. Let this spatial map be called `error map'.
\item Train a Support Vector Machine (SVM) model to classify patches of this error-map image into one of the two categories: new changes present, new changes absent.
\item Once the trained SVM model is obtained, it can be applied to generate binary weights-map for new test images in the same longitudinal study.
\item Once the binary weights-map $\boldsymbol{W_b}$ is obtained, the prior and the pilot can be stitched by $\boldsymbol{W_b}(\rm{Pilot}) + (1-\boldsymbol{W_b})(\rm{Prior})$. Fig.~\ref{fig:okra_svm} shows results of this method on 2D okra dataset.

  Alternatively, these binary weights can also be used in the alternate minimization optimization described in Eq.~4 of the main paper, although this approach has not been pursued here.

  \begin{comment}
\begin{equation}
%   \setlength{\belowdisplayskip}{0pt} \setlength{\belowdisplayshortskip}{0pt}
%\setlength{\abovedisplayskip}{-2pt} \setlength{\abovedisplayshortskip}{-2pt}
J_3(\boldsymbol{\theta},\boldsymbol{\alpha}) = \lVert\boldsymbol{\mathcal{R} x}-\boldsymbol{y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1 +\lambda_2\lVert\boldsymbol{W_b}(\boldsymbol{x} - (\boldsymbol{\mu} + \sum_{k}\boldsymbol{V_k}\alpha_k))\rVert_2^2.
\label{Eq:main}
\end{equation}
\end{comment}
\end{enumerate}

@@

        \caption{}
    \end{subfigure}    
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/svm/okra/resultStitched_svm.png}
        \caption{}
     \end{subfigure}
     \caption[Selecting $k$]{Using an alternate approach to generate (binary) weights map. (a)-(c) the images used as object-prior, (d) the test (310$\times$310). Measurements along $60$ views were taken. (e) error-map showing new regions, (f) detected binary weights map showing regions of new changes, (g) pilot, (h) final reconstruction with pilot in the new regions, and prior in the other regions. } %(h) prior, which is obtained by projecting the pilot onto the eigenspace computed from the templates, 
\label{fig:okra_svm}
\end{figure}
\clearpage

\newpage

@@

\begin{table}[!h]
  \centering
      \caption{SSIM values (within RoI) of reconstructions shown in Figure~\ref{fig:comparisonLit_results}.}
  \begin{tabular}{|l|l|l|}
\hline
 & \textbf{SSIM (whole image)} & \textbf{SSIM (RoI)} \\ \hline
\textbf{Method in~\cite{Lee2012}} & 0.65 & 0.60 \\ \hline
\textbf{\begin{tabular}[c]{@{}l@{}}Weighted prior\\reconstruction\end{tabular}} & 0.89 & 0.92 \\ \hline
  \end{tabular}
  \label{tab:comparisonLit}
\end{table}
\clearpage
\newpage

>>>>

\usepackage{boxedminipage}
%\usepackage{geometry}
\usepackage[colorlinks=true]{hyperref}
\title{Supplemental material for `Eliminating object prior-bias from sparse-projection
tomographic reconstructions'}
\begin{document}
\date{}
\maketitle
This supplementary material consists of details which we did not
include in the main paper for purposes of brevity. This document
consists of links to view our complete 3D reconstructions, an
alternate simpler motivation for the use of multiple eigenspaces, and an
alternate method to compute the parameters for the weights map. We
also include a comparison with a method in literature based on
comments from a reviewer.

\tableofcontents

@@

\section{Links to view our 3D reconstructions}
Here we present the 3D reconstruction results discussed in Sec.~6 of
the main paper. For each dataset, the links contain 2 videos: one
showing all the templates and test volumes, and the other showing the
reconstructed volumes by various methods. The videos are in
\verb+.avi+ format and display reconstructions in a slice-by-slice
manner.

@@

%------------------------------------------------------
\section{Using multiple eigenspaces}

Here we present a schematic explaining the advantage of using multiple
eigenspaces within the proposed spatially-varying reconstruction
technique. This may be viewed as a simpler version of the somewhat
long algorithm described in the main paper.

@@

%\clearpage
%----------------------------------------------------------
\section{Learning-based computation of weights map}

The value of $k$ (discussed in Sec.~7 of the main paper) is typically
dependent on the number of false positives the system can allow in
exchange for the detection of more true positives (regions of new
changes). It is this choice and not the particular type of dataset
that affects the selection of $k$. In our work, the chosen $k$
corresponds to the one that gives minimal tolerance to false
positives. However, in case one wishes to completely avoid the use of
this hyper-parameter, one can construct a binary weights-map using a
method described below.

\begin{enumerate}
\item For the purposes of training, treat one of the previously
  scanned images as test. Annotate the region of differences between
  the test and all of the previously scanned images based on domain
  knowledge.

\item Compute the difference between the pilot reconstruction of the
  test and its projection onto the eigenspace. Let this spatial map be
  called `error map'.

\item Train a Support Vector Machine (SVM) model (or any other
  classifier model) to classify patches of this error-map image into
  one of the two categories: new changes present, new changes absent.

\item Once the trained SVM model is obtained, apply the model and
  generate binary weights-map for new test images in the same
  longitudinal study.

\item Once the binary weights-map $\boldsymbol{W_b}$ is obtained, use
  these in the alternate minimization optimization described in Eq.~7
  of the main paper.
\end{enumerate}

We mention in passing that if we are to entirely avoid the other
parameters (viz. $\lambda_1$ and $\lambda_2$, the prior and the pilot
can be stitched in a simplistic way using the equation 
$\boldsymbol{W_b}(\rm{Pilot}) + (1-\boldsymbol{W_b})(\rm{Prior})$.
The advantage of this method is that the computational time is
signficantly reduced since the alternate minimization procedure is not
used.

%Fig.~\ref{fig:okra_svm} shows results of this method on 2D okra dataset.

  \begin{comment}
\begin{equation}
%   \setlength{\belowdisplayskip}{0pt} \setlength{\belowdisplayshortskip}{0pt}
%\setlength{\abovedisplayskip}{-2pt} \setlength{\abovedisplayshortskip}{-2pt}
J_3(\boldsymbol{\theta},\boldsymbol{\alpha}) = \lVert\boldsymbol{\mathcal{R} x}-\boldsymbol{y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1 +\lambda_2\lVert\boldsymbol{W_b}(\boldsymbol{x} - (\boldsymbol{\mu} + \sum_{k}\boldsymbol{V_k}\alpha_k))\rVert_2^2.
\label{Eq:main}
\end{equation}
\end{comment}
%\end{enumerate}

@@

        \caption{}
    \end{subfigure}    
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/svm/okra/resultStitched_svm.png}
        \caption{}
     \end{subfigure}
     \caption[Selecting $k$]{Using an alternate approach to generate (binary) weights map. (a)-(c) the images used as object-prior, (d) the test (310$\times$310). Measurements along $60$ views were taken. (e) error-map showing new regions, (f) detected binary weights map showing regions of new changes, (g) pilot, (h) final reconstruction with pilot in the new regions, and prior in the other regions. } %(h) prior, which is obtained by projecting the pilot onto the eigenspace computed from the templates, 
\label{fig:okra_svm}
\end{figure}
%\clearpage
%\newpage

@@

\begin{table}[!h]
  \centering
      \caption{SSIM values (within RoI) of reconstructions shown in Figure~\ref{fig:comparisonLit_results}.}
  \begin{tabular}{|l|l|l|}
\hline
 & \textbf{SSIM (whole image)} & \textbf{SSIM (RoI)} \\ \hline
\textbf{Method in~\cite{Lee2012}} & 0.65 & 0.60 \\ \hline
\textbf{\begin{tabular}[c]{@{}l@{}}This paper\end{tabular}} & 0.89 & 0.92 \\ \hline
  \end{tabular}
  \label{tab:comparisonLit}
\end{table}
%\clearpage
%\newpage
